---
title: "ANA 515 Assignment 4 Data Analytics Project"
author: "Mohammed Sohail Ameer"
date: "2022-10-15"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Business Goal

Business problem or goal is a essential part of a business dataset analysis. Here we are utilized Uber's data where we are finding the goals or problems in it. Here the goal is to find out the frequency of a customer travelled in Uber and what type of service they are willing to avail. Again to achieve a good service and customer reputation, the cancelled and completed trips are also analysed with their fare, distance, ride time etc.

## Dataset Retrieval

This Uber's data is retrieved from the following link:

https://drive.google.com/file/d/1emopjfEkTt59jJoBH9L9bSdmlDC4AR87/view

## Intializing required libraries

```{r}
library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(DT)
library(scales)
library(Hmisc)
```

## Loading Data

```{r}
apr_data <- read.csv("Uber/uber-raw-data-apr14.csv")
may_data <- read.csv("Uber/uber-raw-data-may14.csv")
jun_data <- read.csv("Uber/uber-raw-data-jun14.csv")
jul_data <- read.csv("Uber/uber-raw-data-jul14.csv")
aug_data <- read.csv("Uber/uber-raw-data-aug14.csv")
sep_data <- read.csv("Uber/uber-raw-data-sep14.csv")

data_2014 <- rbind(apr_data,may_data, jun_data, jul_data, aug_data, sep_data)
```

## Describing Data

```{r}
describe((data_2014))
```
## Data Preparation

```{r}
data_2014$Date.Time <- as.POSIXct(data_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S")
data_2014$Time <- format(as.POSIXct(data_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S"), format="%H:%M:%S")
data_2014$Date.Time <- ymd_hms(data_2014$Date.Time)
data_2014$day <- factor(day(data_2014$Date.Time))
data_2014$month <- factor(month(data_2014$Date.Time, label = TRUE))
data_2014$year <- factor(year(data_2014$Date.Time))
data_2014$dayofweek <- factor(wday(data_2014$Date.Time, label = TRUE))
data_2014$hour <- factor(hour(hms(data_2014$Time)))
data_2014$minute <- factor(minute(hms(data_2014$Time)))
data_2014$second <- factor(second(hms(data_2014$Time)))
```

## Data Visualization

Here a bar plot has been developed with Trips by Hour and Month. 

```{r}
hour_data <- data_2014 %>%
           group_by(hour) %>%
               dplyr::summarize(Total = n()) 
month_hour <- data_2014 %>%
          group_by(month, hour) %>%
             dplyr::summarize(Total = n())
ggplot(month_hour, aes(hour, Total, fill = month)) + 
       geom_bar( stat = "identity") +
          ggtitle("Trips by Hour and Month") +
           scale_y_continuous(labels = comma)
```
## Data Visualization

Here a heat map plot has been developed with Heat Map by Hour and Day.

```{r}
day_and_hour <- data_2014 %>%
         group_by(day, hour) %>%
            dplyr::summarize(Total = n())
ggplot(day_and_hour, aes(day, hour, fill = Total)) +
            geom_tile(color = "black") +
              ggtitle("Heat Map by Hour and Day")
```

